---
title: "Joins In Class Activity"
format: 
  html:
    code-tools: true
    toc: false
execute: 
  echo: true
---


This in-class file is used for demonstrating individual examples related to this topic and is kept as a persistent record of examples provided during lectures.

## The Data

The data we will be using for this is from a library that has several `data.frame` objects related to flights out of three New York airports

```{r}
library( nycflights13 )
```

The data.frames include four `data.frame` objects with external data (each of which has a *primary key*), including:

```{r}
head(airlines)
```

```{r}
head( airports )
```

```{r}
head( planes )
```

```{r}
head(weather)
```

And one *large* `data.frame` that has columns that are *foreign keys* to those previous `data.frame` objects.

```{r}
names( flights )
```


## Asking Questions

Using these data, we will asks some of the following questions about the data.



> Which NYC airport has the worst departure delays?


```{r}
library( tidyverse )
library( knitr )
flights %>%
  select( origin, dep_delay ) %>% # select origin & dep_delay
  group_by( origin ) %>% # group_by( origin )
  summarise( Maximum = max( dep_delay, na.rm=TRUE  ),   # summarize max origin & value
             Mean = mean( dep_delay, na.rm=TRUE),
             Median = median( dep_delay, na.rm=TRUE) ) %>%
  left_join( airports, by=c("origin"="faa") ) %>%
  select( Airport = name, 
          Mean, Median, Maximum )   %>% # join with airport to get real name
  kable()
# make a table


```



> Which models of aircraft have the longest average flight distance?

```{r}
flights %>%
  select( distance, tailnum ) %>%
  group_by( tailnum) %>%
  summarize( dist = mean( distance, na.rm=TRUE ) ) %>%
  left_join( planes, by="tailnum" ) %>%
  select( model, dist ) %>%
  arrange( -dist ) %>%
  group_by( model ) %>%
  summarize( dist = mean( dist, na.rm=TRUE ) ) %>%
  arrange( -dist )
```





























```{r}
flights %>%
  select( tailnum, distance ) %>%
  group_by( tailnum ) %>%
  summarise( dist = mean( distance, na.rm=TRUE ) ) %>%
  left_join( planes ) %>%
  select( dist, model ) %>%
  group_by( model ) %>%
  summarise( dist = mean( dist ) ) %>%
  arrange( -dist ) %>% 
  head()
```


> Which airlines has the most has the most on-time arrivals?











