---
title: "Joins In Class Activity"
format: 
  html:
    code-tools: true
    toc: false
execute: 
  echo: true
---


This in-class file is used for demonstrating individual examples related to this topic and is kept as a persistent record of examples provided during lectures.

## The Data

The data we will be using for this is from a library that has several `data.frame` objects related to flights out of three New York airports

```{r}
library( nycflights13 )
```

The data.frames include four `data.frame` objects with external data (each of which has a *primary key*), including:

```{r}
head(airlines)
```

```{r}
head( airports )
```

```{r}
head( planes )
```

```{r}
head(weather)
```

And one *large* `data.frame` that has columns that are *foreign keys* to those previous `data.frame` objects.

```{r}
names( flights )
```


## Asking Questions

Using these data, we will asks some of the following questions about the data.



> Which NYC airport has the worst departure delays?



> Which models of aircraft have the largest average flight distance?



> Which airlines has the most has the most on-time arrivals?











